---
title: "Practice 2.2"
author: "Nathanaël LAHARY : T12508101"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Code 

Github reposit : [link](https://github.com/Nathanael-FR/OCEAN5098)

# Objective

This short experiment aims at testing the efficiency of a diet made of bubble teas: three cups a day for a week. Weight of subjects is provided before and after this ‘bubble tea’ diet (the data are fake).

### Creating the dataframe

```{r data}
before_diet <- c(104,95,87,77,112)
after_diet <- c(96,91,81,75,118)
df <- data.frame(before_diet,after_diet) 
row.names(df) <- c("subject_1","subject_2","subject_3","subject_4","subject_5")
df
```

### Import the libraries
```{r libraries, message=FALSE}

library(tidyr)
library(dplyr)
```

### Reformat the data

**STEP 1: **Reformat this data frame to obtain the weight as a double into one column and the time of measurement as a factor with two levels before_diet and after_diet.

```{r pivot}
# Reformat the data
df_reformated <- df %>%
    tibble::rownames_to_column(var = "Subject") %>%
    pivot_longer(cols = c(before_diet,after_diet),
               names_to = "time",
               values_to = "weight")

df_reformated$weight <- as.double(df_reformated$weight)
df_reformated$time <- as.factor(df_reformated$time)
df_reformated
```
 
**STEP 2: **You will store this data frame as the first element of a list called BUBBLE_DIET. The second element of this list will be another list called WEIGHT_LOSS, storing three elements: 
[1] a vector of character extracting the row names of the table previously created; 
[2] a numeric vector (double) indicating the weight loss of each subject (can be positive or negative, but calculated as a percentage of initial weight); 
[3] a combination of these two elements in a table with two columns: subject and weight_loss.

*Row names and Weight loss*
```{r unite}
# Get patients ID
Patients <- row.names(df)

# Get the weight loss for each of them
Weight_loss <- df_reformated %>%
  group_by(Subject) %>%
  summarise(weight_loss = (weight[time == "before_diet"] - weight[time == "after_diet"]) / weight[time == "before_diet"] * 100,
            .groups = 'drop')

Weight_loss <- Weight_loss$weight_loss

Patients
Weight_loss
```

*List*
```{r remove}
BUBBLE_DIET <- list(df_reformated,
                    WEIGHT_LOSS = list(
                      Patients,
                      Weight_loss,
                      data.frame(Patients,Weight_loss)),
                    Message = "I actualy like R, but sometimes I wonder if some operations
                    would be easier to do on a wide df than on a long one. Like the weight_loss%, for me
                    it was more instinctive to get it on the first df. See the example below :"
                    )

# I entirely agree with you, but the aim was to move around with objects extracted from a data frame. Your solution  is indeed more elegant, and I can see it is very easy for you :) Very good. 

BUBBLE_DIET
```
### Full code
```{r fullcode, message=FALSE}
#setwd("C:/Users/natha/Documents/R_NTU_2023/Course 4/Practice 3.2")
rm(list=ls())


before_diet <- c(104,95,87,77,112)
after_diet <- c(96,91,81,75,118)
df <- data.frame(before_diet,after_diet) 
row.names(df) <- c("subject_1","subject_2","subject_3","subject_4","subject_5")

library(dplyr)
library(tidyr)

# Reformat the data
df_reformated <- df %>%
    tibble::rownames_to_column(var = "Subject") %>%
    pivot_longer(cols = c(before_diet,after_diet),
               names_to = "time",
               values_to = "weight")

df_reformated$weight <- as.double(df_reformated$weight)
df_reformated$time <- as.factor(df_reformated$time)

# Check data types

# df_reformated$time
# class(df_reformated$time)
# typeof(df_reformated$time)

#typeof(df_reformated$weight)

# Get patients ID
Patients <- row.names(df)

# Get the weight loss for each of them
Weight_loss <- df_reformated %>%
  group_by(Subject) %>%
  summarise(weight_loss = (weight[time == "before_diet"] - weight[time == "after_diet"]) / weight[time == "before_diet"] * 100,
            .groups = 'drop')

Weight_loss <- Weight_loss$weight_loss

# Create a list to gather everything
BUBBLE_DIET <- list(df_reformated,
                    WEIGHT_LOSS = list(
                      Patients,
                      Weight_loss,
                      data.frame(Patients,Weight_loss)),
                    Message = "I actualy like R, but sometimes I wonder if some operations
                    would be easier to do on a wide df than on a long one. Like the weight_loss%, for me
                    it was more instinctive to get it on the first df. See the example below :"
                    )



# Second approach

# df$before_diet <- as.double(df$before_diet)
# df$after_diet <- as.double(df$after_diet)
# df <- df %>%
#   mutate(Weight_loss = (before_diet - after_diet)/before_diet*100)
#   
# Weight_loss2 <- df$Weight_loss

```
