---
title: "Practice 2.2"
author: "Nathanaël LAHARY : T12508101"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

Datasheet : [link](https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt)

```{r data, echo=FALSE,fig.width=120}
options(width = 120)
rairuoho<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',
                     header=T, 
                     sep="\t")
rairuoho
```

# Code 

Github reposit : [link](https://github.com/Nathanael-FR/OCEAN5098)

### Import the data
```{r data_import}
rairuoho<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',
                     header=T, 
                     sep="\t")
``` 

### Import the libraries
```{r libraries, message=FALSE}

library(tidyr)
library(dplyr)
```

### Data cleaning

**STEP 1: **Format the table in order to have the day as a single variable. Length will be created to receive the length of the grass at the given day.

```{r pivot}
rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length")
rairuoho_formated
```
 
**STEP 2: **Replace nutrient with enriched in the data set. Make day as a single variable (factor) containing 6 levels (day3, day4, day5, day6, day7, day8).

```{r mutate}
rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length") %>%
  mutate(treatment = ifelse(treatment=="nutrient","enriched",treatment),
         day = factor(day,levels = c("day3","day4","day5","day6","day7","day8")))
rairuoho_formated
```

**STEP 3: **Merge variables Spatial1 and Spatial2 in order to obtain the following format for the spatial coordinates of an observation upper_left.

```{r unite}
rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length") %>%
  mutate(treatment = ifelse(treatment=="nutrient","enriched",treatment),
         day = factor(day,levels = c("day3","day4","day5","day6","day7","day8"))) %>%
  unite(c("spatial1","spatial2"),col="Spatial",sep = "_") # Function found in the cheat sheet p17

rairuoho_formated
```

**STEP 4: **Remove variables row and column.
```{r remove}
rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length") %>%
  mutate(treatment = ifelse(treatment=="nutrient","enriched",treatment),
         day = factor(day,levels = c("day3","day4","day5","day6","day7","day8"))) %>%
  unite(c("spatial1","spatial2"),col="Spatial",sep = "_") %>% # Function found in the cheat sheet p17
  select(-row,-column) # mutate(row=NULL,column=NULL) also works

rairuoho_formated
```
### Full code
```{r fullcode}
library(tidyr)
library(dplyr)

rairuoho<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',
                     header=T, 
                     sep="\t")

rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length") %>%
  mutate(treatment = ifelse(treatment=="nutrient","enriched",treatment),
         day = factor(day,levels = c("day3","day4","day5","day6","day7","day8"))) %>%
  unite(c("spatial1","spatial2"),col="Spatial",sep = "_") %>% # Function found in the cheat sheet p17
  select(-row,-column) # mutate(row=NULL,column=NULL) also works


# Another method 
#
# rairuoho$treatment <- ifelse(rairuoho$treatment=="nutrient","enriched","water")
# rairuoho$row = NULL
# rairuoho$column = NULL
# rairuoho_formated_bis <- rairuoho %>%
#   pivot_longer(cols = day3:day8, names_to = "day", values_to = "length") %>%
#   mutate(day = factor(day, levels = c("day3","day4","day5","day6","day7","day8"))) %>%
#   unite(c("spatial1","spatial2"),col="Spatial",sep = "_")
```
### Results
```{r result,echo=FALSE}
library(tidyr)
library(dplyr)

rairuoho<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',
                     header=T, 
                     sep="\t")

rairuoho_formated <- rairuoho %>% 
  pivot_longer(cols=day3:day8,names_to = "day", values_to = "length") %>%
  mutate(treatment = ifelse(treatment=="nutrient","enriched",treatment),
         day = factor(day,levels = c("day3","day4","day5","day6","day7","day8"))) %>%
  unite(c("spatial1","spatial2"),col="Spatial",sep = "_") %>% # Function found in the cheat sheet p17
  select(-row,-column) # mutate(row=NULL,column=NULL) also works

print(rairuoho_formated,n=Inf)
```

VD: perfect. This edit so that you know how to pull my request in your code. 
